---
title: "Co-inoculation experiments"
author: "Tia Harrison"
date: "09/04/2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


## Load the packages for analysis

```{r packages}
# Packages for analysis
library(car) 
library(lsmeans)
library(multcomp)
library(lme4)
library(performance)
library(see)
library(gridExtra)
library(wesanderson)
library(RColorBrewer)
library(R2admb)
library(glmmADMB)
library(dplyr)
library(tidyverse)
library(nationalparkcolors)
library(ggpubr)
```

## Set contrasts 
When performing an Anova, should set contrasts to use effects coding instead of the default (treatment contrasts).

```{r contrasts}
options(contrasts = c("contr.sum","contr.poly")) 
```

## Overdispersion function 
Ben Bolker's function for testing overdispersion in poisson models  https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#overdispersion 

```{r overdisp}
overdisp_fun <- function(model) {
rdf <- df.residual(model)
rp <- residuals(model,type="pearson")
Pearson.chisq <- sum(rp^2)
prat <- Pearson.chisq/rdf
pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```


# Data 

## Load and explore the two datasets for analysis 
The Bacillus data set was collected from the growth chamber experiment where plants that are tropical in origin (Costa Rica) were used in the experiment. The Pseudomonas data set was collected from the greenhouse experiment and the seeds and strains used in that experiment were from a temperate population in Virginia.

```{r data}
bacillus <- read_csv("Bacillus_data.csv") # This data already has the initial dead plants removed from it and has 95 plants in it
pseudomonas <- read_csv("Pseudomonas_data.csv") # This data set is still the full dataset and has 270 plants in it 

# Convert biomass to mg in bacillus dataset 
# Pseudomonas data set is already in mg 
# Convert to factor 
bacillus <- bacillus%>% 
  mutate(Aboveground_mg= Aboveground_g*1000) %>%
  mutate(Belowground_mg=Belowground_g*1000) %>%
  mutate(Rhizobia=as.factor(Rhizobia), 
         Bacillus=as.factor(Bacillus), 
         Block=as.factor(Block), 
         Plant_ID=as.factor(Plant_ID))

# Remove plants that died before treatment took effect 
# Convert relevant data to factors
pseudomonas2 <- pseudomonas %>% 
  filter(Mortality1 == 1) %>%
  mutate(Bradyrhizobia=as.factor(Bradyrhizobia), 
         Pseudomonas=as.factor(Pseudomonas), 
         Variovorax=as.factor(Variovorax),
         Block=as.factor(Block), 
         Plant_ID=as.factor(Plant_ID)) # Now the experiment has 216 plants in it 

# Break down of how many plants alive at the end of the experiment 
bacillus3 <- bacillus %>%
  filter(Dead ==0) # 64 plants but if they are in the treatments with rhizobia might be enough to compare 

pseudomonas3 <- pseudomonas2 %>%
  filter(Dead ==0) # 164 plants alive at the end of the experiment 
```

# Analysis 

## Tropical non-rhizobia strains 
Test if the Bacillus strain have an impact on plant growth, nodule number, and nodule weight belowground, and whether or not there is an interaction with rhizobia inoculations. 

```{r tropical nr_model}
# histogram of aboveground data 
hist(bacillus3$Aboveground_mg) 

# aboveground model 
mod_full_s <- lmer(log10(Aboveground_mg) ~ Bacillus*Rhizobia + (1|Family), data=bacillus3)

# Check assumptions 
plot(mod_full_s)
qqnorm(resid(mod_full_s))
qqline(resid(mod_full_s))
plot(fitted(mod_full_s), sqrt(abs(resid(mod_full_s))), main="Scale-location")
# Tried a bunch of things but the results came out all the same and this is the best fitting model 

# Test significance and look at summary 
summary(mod_full_s)
Anova(mod_full_s, type=3)

# Nodule analysis 
hist(bacillus3$Nodule_no) 

# Include random term for overdispersion 
mod_nod_s <- glmer(Nodule_no ~ Bacillus*Rhizobia + (1|Family) + (1|Plant_ID), family=poisson(link="sqrt"),  data=bacillus3)

# Test fit 
plot(mod_nod_s)
qqnorm(resid(mod_nod_s))
qqline(resid(mod_nod_s))
plot(fitted(mod_nod_s), sqrt(abs(resid(mod_nod_s))), main="Scale-location")

# Check overdispersion 
overdisp_fun(mod_nod_s) # if non-significant then no overdispersion 

# Test significance and look at summary 
summary(mod_nod_s)
Anova(mod_nod_s, type=3)

# Nodule weight 
# Only performed on plants that formed nodules and so we removed Control and treatments without rhizobia
bacillus_r <- bacillus3 %>%
  filter(Rhizobia=="YES")

# Look at the data 
hist(bacillus_r$Nodule_weight_ug) 

# Run model 
mod_weight <- lmer(Nodule_weight_ug ~ Bacillus + (1|Family), data=bacillus_r)

# Test significance and look at summary 
Anova(mod_weight, type=3)
summary(mod_weight)

# Check assumptions 
plot(mod_weight)
qqnorm(resid(mod_weight))
qqline(resid(mod_weight))
plot(fitted(mod_weight), sqrt(abs(resid(mod_weight))), main="Scale-location") # everything looks good 
```

## Temperate non-rhizobia strains 
Test whether the non-rhizobia strains Pseudomonas and Variovorax have impacts on plant growth and nodule formation when coinoculation with rhizobia. Look for interactions between all three strains inoculated on the plants. 

```{r temperate nr_model}
# histogram of aboveground data 
hist(pseudomonas3$Aboveground_mg) 

# Model on full data set with all treatments
mod_full_ps <- lmer(log10(Aboveground_mg) ~ Bradyrhizobia*Pseudomonas*Variovorax + (1|Family) + (1|Block), data=pseudomonas3)

# Check assumptions 
plot(mod_full_ps)
qqnorm(resid(mod_full_ps))
qqline(resid(mod_full_ps))
plot(fitted(mod_full_ps), sqrt(abs(resid(mod_full_ps))), main="Scale-location") # Pretty good fit! 

# Test significance and look at summary 
summary(mod_full_ps)
Anova(mod_full_ps, type=3)


# Nodules 
# Data is overdispersed and zero -inflated but when we tried zero inflated models the model was overparametized
# Ttesting the rhizobia treatments only worked best 
pseudomonas_r <- pseudomonas3 %>%
  filter(Bradyrhizobia=="Y")

# look at the data 
hist(pseudomonas3$Nodule_no) 

# Fit the model 
mod_nod_glm_full <- glmer(Nodule_no ~ Pseudomonas*Variovorax + (1|Family) + (1|Block) + (1|Plant_ID), family=poisson(link="sqrt"), data=pseudomonas_r)

# Check assumptions 
plot(mod_nod_glm_full)
qqnorm(resid(mod_nod_glm_full))
qqline(resid(mod_nod_glm_full))
plot(fitted(mod_nod_glm_full), sqrt(abs(resid(mod_nod_glm_full))), main="Scale-location")
overdisp_fun(mod_nod_glm_full)

# Check significance  
summary(mod_nod_glm_full)
Anova(mod_nod_glm_full, type=3)
```
## Tests of variance 
Testing whether variance in nodule number in the rhiozbia only treatments is different from in coinoculation treatments in both the tropical and temperate experiment data. 

```{r variance}
# tropical experiment 
trop_variance <- bartlett.test(Nodule_no ~ Bacillus, bacillus_r)
trop_variance

# temperate experiment 
temp_variance <- leveneTest(Nodule_no ~ Pseudomonas*Variovorax, pseudomonas_r)
temp_variance
```


# Figures 
Plotting the data for the tropical and temperate datasets 

## Plotting aboveground biomass 
```{r biomass plot}
# Specify the colour palette 
treatments3 <- park_palette("RockyMountains")
grey_col <- brewer.pal(9, "Greys") 
 
# Prep the data 
bacillus34 <- bacillus3 %>%
  mutate(rhizobia_new=ifelse(Rhizobia=="YES", "Rhizobia", "Control")) %>%
  mutate(bacillus_new=ifelse(Bacillus=="YES", "          Bacillus", "Control")) %>%
  mutate(bacillus_new=fct_relevel(bacillus_new, "Control", "          Bacillus"))

# Plot tropical experiment full dataset 
bacillus_mass <- ggplot(bacillus34, aes(x=bacillus_new, y=log10(Aboveground_g), fill=Rhizobia)) + 
  geom_boxplot(fatten=1) +
  (ylab("Log biomass (mg)"))+
  (xlab(""))+
  labs(tag = "(a)") +
  facet_grid(~rhizobia_new, switch="x") +
  scale_fill_manual(values = c("white", grey_col[4])) +
  theme_classic() +
  theme(legend.position="none",
        axis.text.x=element_text(size=9, hjust = 0.98, angle=75),
        strip.background =element_rect(fill=grey_col[2]))

# Prep temperate data 
pseudomonas34 <- pseudomonas3 %>%
  mutate(rhizobia_new=ifelse(Bradyrhizobia=="Y", "Rhizobia", "Control")) %>%
  mutate(treatment_new=ifelse(Pseudomonas =="N" & Variovorax =="N", "Control", Treatment)) %>%
  mutate(treatment_new=ifelse(Pseudomonas=="Y" & Variovorax=="N", "Pseudomonas", treatment_new)) %>%
  mutate(treatment_new=ifelse(Pseudomonas=="N" & Variovorax=="Y", "Variovorax", treatment_new)) %>%
  mutate(treatment_new=ifelse(Pseudomonas=="Y" & Variovorax=="Y", "All strains", treatment_new)) %>%
  mutate(treatment_new=fct_relevel(treatment_new, "Control", "Pseudomonas", "Variovorax", "All strains"))

# Plot the temperate data 
pseud_mass <- ggplot(pseudomonas34, aes(x=treatment_new, y=log10(Aboveground_mg), fill=rhizobia_new)) + 
  geom_boxplot(fatten=1) +
  (ylab("Log biomass (mg)"))+
  (xlab(""))+
  labs(tag = "(b)") +
  facet_wrap(~rhizobia_new, switch="x") +
  scale_fill_manual(values = c("white", grey_col[4])) +
  theme_classic() +
  theme(legend.position="none", 
        axis.text.x=element_text(size=9, hjust = 0.98, angle=75),
        strip.background =element_rect(fill=grey_col[2]))

# Look at the plot 
grid.arrange(bacillus_mass, pseud_mass,  
             ncol=3, nrow=1, 
             layout_matrix=rbind(c(1,2,2)))

# Save the plot 
# pdf(file="Biomass_plot.pdf", width=7, height=5)

test3<-grid.arrange(bacillus_mass, pseud_mass,  
             ncol=3, nrow=1, 
             layout_matrix=rbind(c(1,2,2)))

# dev.off()
ggsave(filename="Figures/Biomass_plot.pdf", plot=test3)
```

## Plotting nodule number 
```{r nodule plot}
# Plot tropical experiment full dataset 
bacillus_nod <- ggplot(bacillus34, aes(x=bacillus_new, y=Nodule_no, fill=Rhizobia)) + 
  geom_boxplot(fatten=1) +
  (ylab("Nodules (no)"))+
  (xlab(""))+
  labs(tag = "(a)") +
  facet_grid(~rhizobia_new, switch="x") +
  scale_fill_manual(values = c("white", grey_col[4])) +
  theme_classic() +
  theme(legend.position="none",
        axis.text.x=element_text(size=9, hjust = 0.98, angle=75),
        strip.background =element_rect(fill=grey_col[2]))

# Plot the temperate data 
pseud_nod <- ggplot(pseudomonas34, aes(x=treatment_new, y=Nodule_no, fill=rhizobia_new)) + 
  geom_boxplot(fatten=1) +
  (ylab("Nodules (no)"))+
  (xlab(""))+
  labs(tag = "(b)") +
  facet_wrap(~rhizobia_new, switch="x") +
  scale_fill_manual(values = c("white", grey_col[4])) +
  theme_classic() +
  theme(legend.position="none", 
        axis.text.x=element_text(size=9, hjust = 0.98, angle=75),
        strip.background =element_rect(fill=grey_col[2]))

# Look at the plot 
grid.arrange(bacillus_nod, pseud_nod,  
             ncol=3, nrow=1, 
             layout_matrix=rbind(c(1,2,2)))

# Save the plot 
# pdf(file="Nodule_plot.pdf", width=7, height=5)

test4<-grid.arrange(bacillus_nod, pseud_nod,  
             ncol=3, nrow=1, 
             layout_matrix=rbind(c(1,2,2)))

#dev.off() 
ggsave(filename="Figures/Nodule_plot.pdf", plot=test4)
```

